\sysfull

### 2018-08-07 paper/asplos19/coalescing.tex
coalescing revised

When a device's buffered energy is sufficient to run multiple tasks without a power failure, committing state after each task is an unnecessary overhead. \sys eliminates this overhead by coalescing a sequence of tasks and deferring the commit
operations for all tasks to the end of the sequence. In general, committing involves moving state manipulated by a task into its permanent location in non-volatile memory. In particular, \sys's commit procedure copies dirty pages of memory that a task updated between fast, volatile working memory and slower, non-volatile main memory. We defer the details of paging privatization and commit to Section~\ref{sec:memory_virtulaization}.
An effective coalescing strategy must be {\em aggressive} enough, attempting to coalesce a large number of tasks to amortize commit overhead.
However, it must also be  {\em conservative} enough, attempting to coalesce only as many tasks as will execute to completion given certain energy conditions, reducing the risk or long coalesced task re-execution penalty.

@@

\begin{algorithm}[t]
	\caption{Coalescing}
	\label{algo:genCoalescing}
	\scriptsize
	\begin{algorithmic}[1]
        \State $C \leftarrow $ \Call{$f_\text{reboot}$}{$z$} \Comment{Update coalescing target by reboot update function}
        \State $H \gets 0$ \Comment{Initialize history: total number of tasks executed since last power failure}
        \While{$\texttt{true}$}       
	        \State $ i \gets 0$
	        %
	        \While{$ i <= C$}
		        \State \Call{execute\_task}{$T_\texttt{i}$}
		        \State $W \leftarrow $ \Call{$f_\text{weight}$}{$T_\text{i}$} \Comment{Assign new task-dependent weight}
		        \State $i \gets i + W$
				\State $H \gets H + W$
	        \EndWhile
	        %
	        \State \Call{commit\_to\_fram}{\null}
	        \State $C \leftarrow $ \Call{$f_\text{commit}$}{$C$}
        \EndWhile
	\end{algorithmic}
\end{algorithm}

@@

Different coalescing strategies adhere mainly to the template in Algorithm~\ref{algo:genCoalescing}, varying in only a few {\em characteristic operations} that the algorithm leaves deliberately abstract. A strategy's reboot update function, $f_\text{reboot}$, updates $C$, the coalescing target, after a reboot. A strategy's weight lookup function, $f_\text{weight}$, returns a task's weight.  A strategy's commit update function, $f_\text{commit}$, updates $C$ after a successful commit. The following paragraphs detail several possible coalescing strategies that we developed for \sys, albeit not the only possible. In fact, \sys \emph{allows programmers to design their own strategy} to implement and link to \sys's task scheduler.

@@

A very simple {\em energy-oblivious} (EO) coalescing strategy treats all tasks as having unity weight and varies the size of a coalesced task linearly. In such a scheme, the number of tasks to coalesce, $C$, increases by a constant $x$ when a coalesced sequence of tasks commits, and decreases by the same constant $x$ when a coalesced sequence of tasks fails to complete, i.e., after a power failure. The characteristic operations of EO are
%
\begin{equation}
	\begin{split}
		f_\text{reboot}(C) & = C - x, \\
		f_\text{weight}(T_\text{i}) & =  1, \\
		f_\text{commit}(C) & = C + x.
	\end{split}
	\label{eq:eo}
\end{equation}
%
The energy-oblivious strategy reacts slowly to the variation in the energy required to execute different tasks and to the variation in the effective quantum of energy available to the device. With the successful commit of each coalesced task sequence, the coalescing target increases. Eventually, the target may be too high and only a coalesced task composed of a few units will commit without interruption by a power failure. The strategy then linearly decreases $C$, eventually reaching a value that allows completion. A key limitation of this algorithm lies in the equality of the target decrease in $f_\text{reboot}$ and the increase in $f_\text{commit}$. Let us assume $x=1$. If, after $k$ successful commits, the target must decrease to its original value $C-k$ due to an energy drop, EO requires $k$ successive power failures, before being able to progress.

@@

The \emph{Energy-Guided} (EG) coalescing strategy adapts its coalescing target more quickly, addressing a key limitation of the EO, to adhere to changes in energy conditions. It uses its recent execution history, H, to estimate energy availability and alters its target accordingly. The EG algorithm is characterized by the following functions
%
\begin{equation}
	\begin{split}
		 f_\text{reboot}(H) & = \lceil \rho H \rceil,\\
		 f_\text{weight}(T_\text{i}) & = 1, \\
		 f_\text{commit}(C) & = \lceil \gamma C \rceil,
	\end{split}
	\label{eq:eg}
\end{equation}
%
where $\rho, \gamma \in [0, 1]$. By relying on the history of execution EG eliminates the problem of frequent power failures on a single coalesced task. In fact, at any reboot, EG will {\em conservatively} try to coalesce only a fraction ($\rho$) of tasks that had successfully completed during the last power cycle.
% In fact EG can reduce its target to one from any coalesced target in at most two successive power failures. For instance, if the execution is interrupted after finishing a single task the history will be one ($H = 1$) and the target will be set to one also. While a history of $H$ suggests that there is sufficient energy to coalesce and successfully commit $H$ tasks (i.e., to set $C = H$). EG conservatively sets $C$ equal to half of $H$ moderating the risk of a very long coalesced task. We experimented with a range of fraction values, settling on one half as the default for EG, because one half had the highest performance.
Figure~\ref{fig:hg-coal} illustrates a snapshot of the operation of EG, with $\rho = \gamma = 0.5$. The figure shows two power cycles following the $i$th, the latter having an execution history $H^{(i)} = 4$. Based on that, $C$ is initially set to two and then decreases to one. Once the value of $C$ reaches one, EG is, in effect, expecting a power failure, justifying the conservative approach. After the power failure, EG uses the most recent history $H^{(i+1)} = 8$ to set $C = 4$ and continue execution.

@@

The \emph{Weighted Energy-Guided} (WEG) coalescing strategy accounts for non-uniform energy and time costs of a program's tasks when setting $C$. Each different task in the program consumes a different amount of energy to execute to completion. The EO and EG coalescing strategies assume that each task has the same cost: for these strategies, $C$ simply corresponds to a target {\em count} of tasks to coalesce, regardless of the individual cost of each task. However, if one task executes for ten seconds and another executes for one second, counting tasks misjudges the amount of {\em work} in the coalesced tasks (and the history).
%
Instead, WEG associates a non-unity weight with each task in the program, and tracks the sum of the weights of tasks in a coalesced sequence. When the sum of weights reaches $C$, the target, WEG commits the coalesced task.
% WEG differs from EO and HG by explicitly accounting for the variation in energy cost of a program's tasks, eliminating the assumption that all tasks do the same work, which is a key limitation.
WEG is characterized the same way as EG (equation (\ref{eq:eg})), except that $f_\text{weight}(T_\text{i}) = W_\text{i}$.
%
% The effectiveness of the WEG hinges on correctly identifying the weight of each task, which WEG assumes is {\em statically} available. Profiling the time and energy cost of tasks in a program is a difficult, orthogonal problem~\cite{cleancut_2018,baghsorkhi_cgo_2018}. WEG could use the result of an arbitrarily sophisticated profiling procedure. To produce a concrete result in this paper, we give WEG access to a simple profile of task run time collected offline using a single fixed input. WEG stores the profile in a lookup table that maps a task's identifier to its weight, making the information available to \sys's scheduler at run time.

>>>>

When a device's buffered energy is sufficient to run multiple tasks without a power failure, committing state after each task is unnecessary overhead. \sys reduces this overhead by coalescing a sequence of tasks and deferring commit
operations for all tasks to the end of the sequence. In general, committing involves moving state manipulated by a task into its permanent location in non-volatile memory. In particular, \sys's commit procedure copies dirty pages of memory that a task updated from fast, volatile working memory to slower, non-volatile main memory. We defer the details of paging privatization and commit to Section~\ref{sec:memory_virtulaization}.
An effective coalescing strategy must be {\em aggressive} enough, attempting to coalesce a large number of tasks to amortize commit overhead.
However, it must also be  {\em conservative} enough, coalescing only as many tasks as will execute to completion given certain energy conditions, reducing the risk of re-execution penalty for long coalesced tasks.

@@

\begin{algorithm}[t]
	\caption{Coalescing}
	\label{algo:genCoalescing}
	\scriptsize
	\begin{algorithmic}[1]
        \State $C \leftarrow $ \Call{$f_\text{reboot}$}{$z$} \Comment{Update coalescing target by reboot update function}
        \State $H \gets 0$ \Comment{Initialize history}
        \While{$\texttt{true}$}       
	        \State $ i \gets 0$
	        %
	        \While{$ i <= C$}
		        \State \Call{execute\_task}{$T_\texttt{i}$}
		        \State $W \leftarrow $ \Call{$f_\text{weight}$}{$T_\text{i}$} \Comment{Assign new task-dependent weight}
		        \State $i \gets i + W$
				\State $H \gets H + W$
	        \EndWhile
	        %
	        \State \Call{commit\_to\_fram}{\null}
                \State $C \leftarrow $ \Call{$f_\text{commit}$}{$C$} \Comment{Update coalescing target by commit update function}
        \EndWhile
	\end{algorithmic}
\end{algorithm}

@@

Different coalescing strategies adhere mainly to the template in Algorithm~\ref{algo:genCoalescing}, varying in only a few {\em characteristic operations} that the algorithm leaves deliberately abstract. The reboot update function, $f_\text{reboot}$, updates $C$, the coalescing target, after a reboot. The weight lookup function $f_\text{weight}$ returns a task's weight. The commit update function $f_\text{commit}$ updates $C$ after a successful commit. The following paragraphs detail three coalescing strategies that we developed for \sys, albeit not the only possible. In fact, \sys \emph{allows programmers to design their own strategy} to implement and link against \sys's task scheduler.

@@

A very simple {\em energy-oblivious} (EO) coalescing strategy treats all tasks as having unity weight and varies the size of the coalescing target linearly. In such a scheme, the number of tasks to coalesce, $C$, increases by a constant $x$ when a coalesced sequence of tasks commits, and decreases by the same constant $x$ when a coalesced sequence of tasks fails to complete, i.e., after a power failure. The characteristic operations of EO are
%
\begin{equation}
	\begin{split}
		f_\text{reboot}(C) & = C - x, \\
		f_\text{weight}(T_\text{i}) & =  1, \\
		f_\text{commit}(C) & = C + x.
	\end{split}
	\label{eq:eo}
\end{equation}
%
EO reacts slowly to the variation in the energy required to execute different tasks and to the variation in the effective quantum of energy available to the device. With the successful commit of each coalesced task sequence, $C$ increases. Eventually, the target may be too high and only a coalesced task composed of a few units will commit without interruption by a power failure. The strategy then linearly decreases $C$, eventually reaching a value that allows completion. A key limitation of this algorithm lies in the equality of the target decrease in $f_\text{reboot}$ and the increase in $f_\text{commit}$. Let us assume $x=1$. If, after $k$ successful commits, the target must decrease to its original value $C-k$ due to an energy drop, EO requires $k$ successive power failures, before being able to progress.

@@

The \emph{Energy-Guided} (EG) coalescing strategy adapts its coalescing target more quickly, addressing a key limitation of the EO, to adhere to changes in energy conditions. It uses its recent execution history, H, to estimate energy availability and alters its target accordingly. The EG algorithm is characterized by the following functions
%
\begin{equation}
	\begin{split}
		 f_\text{reboot}(H) & = \lceil \rho H \rceil,\\
		 f_\text{weight}(T_\text{i}) & = 1, \\
		 f_\text{commit}(C) & = \lceil \gamma C \rceil,
	\end{split}
	\label{eq:eg}
\end{equation}
%
where $\rho, \gamma \in [0, 1]$. By relying on the history of execution EG eliminates the problem of frequent power failures on a single coalesced task. In fact, at each reboot, EG will {\em conservatively} try to coalesce only a fraction ($\rho$) of tasks that had successfully completed during the last power cycle.
% In fact EG can reduce its target to one from any coalesced target in at most two successive power failures. For instance, if the execution is interrupted after finishing a single task the history will be one ($H = 1$) and the target will be set to one also. While a history of $H$ suggests that there is sufficient energy to coalesce and successfully commit $H$ tasks (i.e., to set $C = H$). EG conservatively sets $C$ equal to half of $H$ moderating the risk of a very long coalesced task. We experimented with a range of fraction values, settling on one half as the default for EG, because one half had the highest performance.
Fig.~\ref{fig:hg-coal} illustrates a snapshot of the operation of EG, with $\rho = \gamma = 0.5$. The figure shows two power cycles following the $i$th, the latter having an execution history $H^{(i)} = 4$. Based on that, $C$ is initially set to two and then decreases to one. Once the value of $C$ reaches one, EG is expecting a power failure, justifying the conservative approach. After the reboot, EG uses the most recent history $H^{(i+1)} = 8$ to set $C = 4$ and continue execution.

@@

The \emph{Weighted Energy-Guided} (WEG) coalescing strategy accounts for non-uniform energy and time costs of a program's tasks when setting $C$. Each different task in the program consumes a different amount of energy to run to completion. The EO and EG coalescing strategies assume that each task has the same cost: for these strategies, $C$ simply corresponds to a target {\em count} of tasks to coalesce, regardless of the individual cost of each task. However, if one task executes for ten seconds and another executes for one second, counting tasks misjudges the amount of {\em work} in the coalesced tasks (and the history).
%
Instead, WEG associates a non-unity weight with each task in the program, and tracks the sum of the weights of tasks in a coalesced sequence. When the sum of weights reaches $C$, the target, WEG commits the coalesced task.
% WEG differs from EO and HG by explicitly accounting for the variation in energy cost of a program's tasks, eliminating the assumption that all tasks do the same work, which is a key limitation.
WEG is characterized the same way as EG (equation (\ref{eq:eg})), except that $f_\text{weight}(T_\text{i}) = W_\text{i}$.
%
% The effectiveness of the WEG hinges on correctly identifying the weight of each task, which WEG assumes is {\em statically} available. Profiling the time and energy cost of tasks in a program is a difficult, orthogonal problem~\cite{cleancut_2018,baghsorkhi_cgo_2018}. WEG could use the result of an arbitrarily sophisticated profiling procedure. To produce a concrete result in this paper, we give WEG access to a simple profile of task run time collected offline using a single fixed input. WEG stores the profile in a lookup table that maps a task's identifier to its weight, making the information available to \sys's scheduler at run time.

### 2018-08-07 paper/asplos19/partialcommit.tex
coalescing revised

\subsection{Task Downscaling}
\label{sec:task_downsizing}

\sys uses {\em task downscaling} to make progress through a task that is too large to complete using the energy that the device can buffer. Task downscaling executes part of the long task and interrupts its execution with a {\em partial commit}, after which the long task continues. If power fails, the partial commit preserves the progress through the task, and execution resumes from the point of the partial commit, rather than from the start of the task. Eventually, after some number of partial commits and power failures, the task will complete and execution will proceed. Task downscaling violates task atomicity in a non-terminating execution, favoring continued progress over atomicity. If a program requires a task's atomicity to be preserved, they can disable task downscaling for that task.

The key design issue for task downscaling is deciding when to partially commit. \sys only applies task downscaling to a task that is likely to be non-terminating.  A task is likely to be non-terminating if it fails to execute completely twice consecutively. The second incomplete execution executes after a power failure, when the device will have fully recharged its energy buffer. If a task cannot complete when executing with a full energy buffer, the task is marked as non-terminating.

% After identifying a task as likely non-terminating, \sys must decide when during the task's execution to partially commit the task's state. \sys sets a timer at the start of the likely non-terminating task, initializing it to a very large value (e.g., an estimate of the maximum execution time using the device's energy buffer). If the timer expires before power fails, \sys partially commits, retaining  the timer value to use for future partial commits. If the timer does not expire before power fails, \sys halves the timer and tries again. The exponential decrease of the timer value converges rapidly to a usable one.

>>>>

% After identifying a task as likely non-terminating, \sys must decide when during the task's execution to partially commit the task's state. \sys sets a timer at the start of the likely non-terminating task, initializing it to a very large value (e.g., an estimate of the maximum execution time using the device's energy buffer). If the timer expires before power fails, \sys partially commits, retaining  the timer value to use for future partial commits. If the timer does not expire before power fails, \sys halves the timer and tries again. The exponential decrease of the timer value converges rapidly to a usable one.
\label{sec:task_downsizing}
\subsection{Task Downscaling}
\sys uses {\em task downscaling} to make progress through a task that is too large to complete using the energy that the device can buffer. Task downscaling executes part of the long task and interrupts its execution with a {\em partial commit}, after which the long task continues. If power fails, the partial commit preserves the progress through the task, and execution resumes from the point of the partial commit, rather than from the start of the task. Eventually, after some number of partial commits and power failures, the task will complete and execution will proceed. Task downscaling violates task atomicity in a non-terminating execution, favoring continued progress over atomicity. If a programmer requires a task's atomicity to be preserved, they can disable task downscaling for that task.
The key design issue for task downscaling is deciding when to partially commit. \sys only applies task downscaling to a task that is likely to be non-terminating.  A task is likely to be non-terminating if it fails to run to completion twice consecutively. The second incomplete run executes after a power failure, when the device will have fully recharged its energy buffer. If a task cannot complete when executing with a full energy buffer, the task is marked as non-terminating.
